### 2023-03-24 reachable-resilience.tex
Up

\mathieu{remarque si on enlève la strong monotony mais en gardant with effective basis of $\uparrow post^*$:}

\begin{proposition}(à prouver)
 leur algorithme for {\sc bounded resilience} termine mais pas correct for WSTS with upward compatibility, not necessarly strong, and an effective basis of $\uparrow post^*$ ?
\end{proposition}

\begin{proof}
leur algorithme termine mais pas correct car l'intersection de post et J car alors le lemmme xxxx est faux (contre-exemple)
\mathieu{

@@

}

>>>>

Keeping the $\uparrow post^*$ effectiveness hypothesis but loosening the strong compatibility one still yields some decidability result for the general reachable-resilience problem.

However when removing strong compatibility, precision is lost and the algorithm do not work anymore.

\begin{proposition}
The algorithm from \cite{DBLP:conf/gg/Ozkan22} for {\sc $k$-reachable-resilience} terminate but is incorrect for WSTS with upward compatibility (not strong) and an effective basis of $\uparrow \post^*$.
\end{proposition}

\begin{proof}{Sketch}
Let us consider a LCS where erasing a message take one step. Then the upward compatibility is not strong. 
For instance with a rule $\delta(q,b) = \text{ add some }b $,
we have $(q,b) \to (q,bb)$, $(q, baaa) \geq (q,b)$ and
$(q, baaa) \to^* (q,bb)$ however it takes more than a single step.

In this case then $\pred(\uparrow \{(q,b)\})$ is not upward-closed. Hence 
$post^* (s) \cap D \subseteq \pred(\uparrow \{(q,b)\})$
do not imply
$\uparrow post^* (s) \cap D \subseteq \pred(\uparrow \{(q,b)\})$ for a downward-closed set $D$.
In particular by lossiness $(q, \epsilon)$ would belong to
$post^* (s)$,
hence $(q,a)$, $(q,aa)$, etc. would belong to $\uparrow post^* (s)$ although not necessarily to
$post^* (s)$. $(q,aaa)$ not being in $pred(\uparrow \{ (q,b)\})$ means the algorithm 
from \cite{DBLP:conf/gg/Ozkan22} would deduce that $1$-rechable-resilience does not hold in this example when $\Safe=\uparrow \{(q,b)\}$ and $\Bad =S$,
because $\uparrow post^* (s) \cap S \not\subseteq \pred(\uparrow \{(q,b)\})$,
despite having 
$post^* (s) \cap S \subseteq \pred(\uparrow \{(q,b)\})$.

\iffalse
\mathieu{

@@

}
\fi

### 2023-03-24 resilience.tex
Up

These three reachability problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. 
So we restrict our framework to the class of infinite-state WSTS. Since most of decidable properties in WSTS are given as effective upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets \Safe and \Bad.
\alain{on peut penser à des ensembles \Bad definis dans une logique booleennne sur les clos par le bas, haut, +...}

For instance, the well-known mutual exclusion property is often modelized in a $d$-counters machine by the property that a counter $c_{mutex}$ must be bounded by (usually) one. Then, the set $\Safe =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $\Bad =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of \Safe. In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that \Bad is always downward closed and \Safe is always upward closed.
%		
RP is decidable for lossy counter machines (LCM) with \Safe and \Bad semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}. In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
%
In what follows, we will always suppose that a downward closed set $D$ is given by its finite set of ideals $I_D=\{J_1, J_2,...,J_n\}$ satisfying $D=\downarrow D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ and that an upward closed set $U$ is given by its finite set $Min_U$ of its minimal elements satisfying $U=\uparrow Min_U$.

Let $U,V$ be two upward-closed sets given by its set of minimal element and $D,E$ be two downward-closed sets given by its ideal decomposition. All the four  inclusions $U \subseteq V$,  $U \subseteq D$, $D \subseteq U$ and $D \subseteq E$ are decidable provided that the ordering is decidable and sets are recursive.
%
\alain{on peut penser à des ensembles \Bad definis dans une logique booleennne sur les clos par le bas, haut, +...}

@@

%%%%%%%%
\subsection{Case: $\Safe=\downarrow \Safe$ and $\Bad=\uparrow \Bad$.}
%
%		cas exclusion mutuelle
%
je prends les minimaux de $\Bad$ et les idéaux de $\Safe$, pas sur que ce soit decidable....$pred(\Safe)$

If $\Safe = \downarrow 0$ and $\Bad = \uparrow s$ . . .

Then resilience kinds of becomes "can we reach $0$ from $\uparrow s$?".

In a reset petri-net, if $\Bad = \uparrow \mu$ for some $\mu \neq 0$,
and $\Safe = \downarrow 0$,
then the resilience problem is similar to the single-state zero-reachability problem, which is undecidable.

So, the idea is that the question "is $0$ reachable from token assigment $\mu$" is undecidable, but here the question is slightly different because we start from $\uparrow \mu$ instead.

>>>>

These three reachability problems are decidable for finite transition systems but undecidable for (general) infinite-state transition systems. 
So we restrict our framework to the class of infinite-state WSTS. Since most of decidable properties in WSTS are given as effective upward or downward closed sets \cite{DBLP:journals/iandc/AbdullaCJT00, DBLP:journals/tcs/FinkelS01}, we consider upward closed or downward closed sets $\Safe$ and $\Bad$.
\alain{on peut penser à des ensembles $\Bad$ definis dans une logique booleennne sur les clos par le bas, haut, +...}

For instance, the well-known mutual exclusion property is often modelized in a $d$-counters machine by the property that a counter $c_{mutex}$ must be bounded by (usually) one. Then, the set $\Safe =  \{c_{mutex} \leq 1\} \times \mathbb{N}^{d-1}$ is downward closed and $\Bad =\{c_{mutex} \geq 2\} \times  \mathbb{N}^{d-1} $ is the upward closed complementary of $\Safe$. In \cite{DBLP:conf/gg/Ozkan22}, the authors considered that $\Bad$ is always downward closed and $\Safe$ is always upward closed.
%		
RP is decidable for lossy counter machines (LCM) with $\Safe$ and $\Bad$ semilinear sets as a consequence of results in Section 3.4 of \cite{DBLP:conf/rp/Schnoebelen10}. In particular, the existing proofs of resilience use the decidability of reachability; but we wish to decide resilience for models with undecidable reachability.
%
In what follows, we will always suppose that a downward closed set $D$ is given by its finite set of ideals $I_D=\{J_1, J_2,...,J_n\}$ satisfying $D=\downarrow D = \downarrow (J_1 \cup J_2 \cup..\cup J_n)$ and that an upward closed set $U$ is given by its finite set $Min_U$ of its minimal elements satisfying $U=\uparrow Min_U$.

Let $U,V$ be two upward-closed sets given by its set of minimal element and $D,E$ be two downward-closed sets given by its ideal decomposition. All the four  inclusions $U \subseteq V$,  $U \subseteq D$, $D \subseteq U$ and $D \subseteq E$ are decidable provided that the ordering is decidable and sets are recursive.

@@

%%%%%%%%
\subsection{Case: $\Safe=\downarrow \Safe$ and $\Bad=\uparrow \Bad$.}
%
%		cas exclusion mutuelle
%

\begin{theorem}\label{down-up}
\textcolor{red}{
The resilience problem is undecidable for WSTS with more than one minimal element, 
$\Safe=\downarrow \Safe$
and $\Bad=\uparrow \Bad$.
}
\end{theorem}

\begin{proof}
TBD
\end{proof}
